Supplementary Information for: Crosslinguistic Word Orders Enable an Efficient Tradeoff between Memory and Surprisal

### 2020-02-07 writeup/si.tex
code

Here we show that our information-theoretic analysis is compatible with models placing the main bottleneck in the difficulty of retrieval.
We describe an information-theoretic formalization of a model that has both a limited WM and an unlimited STM.

We consider a model that maintains both a small working memory $m_t$ and an unlimited short-term memory $s_t$.

Predictions are made based on working memory $m_t$, incurring surprisal $H[w_t|m_t]$.
In each time step, there is some amount of communication between $m_t$ and $s_t$, corresponding to retrieval operations.
We model this using a variable $r_t$ representing the information that is retrieved from $s_t$.

@@

$m_t = f(x_t, m_{t-1}, r_t)$

$r_t = g(m_{t-1}, s_{t-1})$

$s_t = h(m_{t-1}, s_{t-1})$

@@

>>>>

Here we show that our information-theoretic analysis is compatible with models placing the main bottleneck in the difficulty of retrieval.
We describe an information-theoretic formalization of a model that has both a limited WM and an unlimited STM.
We consider a model that maintains both a small working memory $m_t$ and an unlimited short-term memory $s_t$.
Predictions are made based on working memory $m_t$, incurring surprisal $H[w_t|m_t]$.
In each time step, there is some amount of communication between $m_t$ and $s_t$, corresponding to retrieval operations.
We model this using a variable $r_t$ representing the information that is retrieved from $s_t$.
Formally, the memory state is determined not just by the input $x_t$ and the previous memory state $m_{t-1}$, but also by the retrieved information:
\begin{equation}
	m_t = f(x_t, m_{t-1}, r_t) 
\end{equation}
The retrieval operation is jointly determined by working memory and short-term memory:
\begin{equation}\label{eq:rt}
	r_t = g(m_{t-1}, s_{t-1}) 
\end{equation}
Finally, the short-term memory can incorporate information from the last word and the working memory:
\begin{equation}
	s_t = h(x_{t-1}, m_{t-1}, s_{t-1}) 
\end{equation}

In the model of \cite{lewis-activation-based-2005}, the working memory corresponds to the \emph{buffers}, which together hold a small and fixed number of individual chunks.
The short-term memory contains an unbounded number of chunks.

In ACT-R, each retrieval operation is initiated through a query constructed based on the current state of the buffers; it returns a chunk from short-term memory that is then placed into a buffer.
In our formalization, $r_t$ reflects the totality of all retrieval operations that are made during the processing of $x_{t-1}$; they happen after $x_{t-1}$ has been observed but before $x_t$ has.
Equation~\ref{eq:rt} reflects that these retrieved chunks are determined by working memory and short-term memory.

In ACT-R, the time it takes to process a word is determined primarily by the time spent retrieving chunks, which is determined by the number of retrieval operations and the time it takes to complete each retrieval operation.
A retrieval operation takes longer if more chunks are similar to the retrieval cue.

If the information content of each chunk is bounded, then a bound on $H[r_t]$ corresponds to a bound on the number of retrieval operations.

The ACT-R model of \cite{lewis-activation-based-2005} does not have an explicit surprisal cost.
Instead, surprisal effects are interpreted as arising because, in less constraining contexts, the parser is more likely to

@@

@@

\section{Example where window model is not optimal}
